#Description
R scripts used to generate figures and tables in paper "Spatial Positioning of Immune Hotspots Reflects the Interplay between B and T Cells in Lung Squamous Cell Carcinoma"
#clear environment
```{r}
rm(list = ls())
```
#libraries
```{r}
library(dplyr)
library(reshape2)
library(ggpubr)
library(ggplot2)
library(survival)
library(survminer)
# library(tidyverse)
library(RTCGA)
library(RTCGA.clinical)
library(gridExtra)
library(RColorBrewer)
library(grid)
library(ggcorrplot)
library(EDASeq)
library(TCGAbiolinks)
library(pheatmap)

```
#load data
```{r}
setwd(dirname(rstudioapi::getSourceEditorContext()$path))
#TCGA
load(file.path("data", "TCGA", "survival_LUSC_LUAD.RData"))
load(file.path("data", "TCGA", "TCGA_LUAD-LUSC_path_EST_ABS_BoLi_DAV.RData"))
load(file.path("data", "TCGA", "TCGA_LUSC-biolinksGeneExp_immuneHotspot.RData"))
load(file.path("data", "TCGA", "TCGA_LUAD-biolinksGeneExp_immuneHotspot.RData"))
load(file.path("data", "TCGA", "Bcell_score_LUSC.RData"))
load(file.path("data", "TCGA", "lusc_mutload.RData"))

#External validation cohort
load(file.path("data", "validation", "validation.RData"))

#set save dir
save_dir <- file.path("results")
dir.create(save_dir)

```

# Data dictionary
## TCGA cohort
## 1. survival_LUSC_LUAD.RData
TCGA LUSC and LUAD datasets includes immune hotspot, %lymphocytes, %cancer, survival data, smoking history, demographic information, disease stages, tumor mutation burden (TMB), and clonal neoantigens

+ *stromal_per, lym_per, tumour_per*: percentages of stromal cells, lymphocytes and tumour cells in total cells automatically detected in H&E slides.
+ *fi*: S_intra/immune, fraction of intratumoral immune hotspots in the immune hotspots.
+ *fc*: S_intra/cancer, proportion of intratumoral immune hotspots to intratumoral immune hotspots + cancer hotspots.
+ *ft*: S_intra/tissue, fraction of ntratumoral immune hotspots in the whole tissue.

## 2. TCGA_LUAD-LUSC_path_EST_ABS_BoLi_DAV.RData
corresponding scores for the same LUAD and LUSC cases with the following datasets:
+ TCGA pathology manual scores, 
+ ESTIMATE, 
+ Davoli, 
+ BoLi
+ CIBERSORT absolute. 

## 3. TCGA_LUSC-biolinksGeneExp_immuneHotspot.RData, TCGA_LUAD-biolinksGeneExp_immuneHotspot.RData
summary of gene expression analysis/enrichment for TCGA LUSC, LUAD up/down regulated immune genes~immune hotspot groups.

## 4. Bcell_score_LUSC.RData
* *LUSCmean_B_cell* B cell scores for LUSC cases estimated by seven algorithms:
+ TIMER
+ CIBERSORT absolute
+ MCP-counter
+ quanTIseq
+ EPIC
+ xCell
+ Danaher

## 5. lusc_mutload.RData
Tumor mutation burden computed as somatic mutations per Mb using maftools

## Validation cohort
## 6. sum_valid
Validation LUSC cohort with serial mIHC sections (n=30 for 10 patients)
+ *x.s, y.s*: coordinates of bottom-right corner of 50x50 μm^2 grids in the slide.
+ *cell.count.all*: number of all cells derived from the registered H&E slide.
+ *cell.count.c*: number of cancer cells derived from the registered H&E slide.
+ *cell.count.l*: number of lymphocytes derived from the registered H&E slide.
+ *Hotspots*: Cancer: cancer hotspots, Cancer-immune: intratumoral immune hotspots, Immune: peritumoral immune hotspots
+ *Region*: TLS: tertiary lymphoid structures, LAG: non-TLS lymphoid aggregates, UD: other regions
+ *T cell subsets* include: cd4: CD4+FOXP3- T cells; cd8: CD8+ T cells; foxp3: CD4+FOXP3+ T cells
+ *B cell subsets* include: cd20: CD20+CXCR5- B cells; cd20cxcr5: CD20+CXCR5+ B cells; cd79bCoexp: CD79b+ B cells; p40: P40+ cancer cells
+ *uc, hem, cxcr5*: stain-negative cells on T and B cell panel and CXCR5 single positive cells 

## 7. df_inter_prop
Proportions of cell-cell interactions in individual immune hotspots
+ *clump_id*: id of individual immune hotspots
+ *hs_dig*: numeric representation of hotspot types

## 8. rtre
Target Registration Error between registered and original slides.
+ *sum_d*: sum of distances between benchmarks on the registered and fixed slide.
+ *mean_rTRE, SD_rTRE*: mean and standard diviation of the Target Registration Error representing which represents the distance between original and transformed points normalized by the diagonal length of an image.

# Functions
```{r}
#Stratify patient groups by variables
stratify_df <- function(df, class_column, variable, res.cutLUSCd){
  df[,class_column] <- df[,variable]
  df[df[,class_column] > res.cutLUSCd[[variable]]$estimate[[1]], class_column] <- "High"
  df[df[,class_column] <= res.cutLUSCd[[variable]]$estimate[[1]], class_column] <- "Low"
  df[,class_column] <- factor(as.character(df[,class_column]), levels = c("Low", "High"))
  return(df)
}


plt_hs_celltype <- function(df, compare_hs, cell_type = 'cd8', cell_name = 'CD8+ T cells'){
  p <- ggplot(subset(df, variable == cell_type), aes(x= Hotspots, y = value)) +
    geom_boxplot(outlier.shape = NA, aes(fill = Hotspots))+
    geom_dotplot(binaxis='y', stackdir='center', stackratio = 0.8,
                 position=position_jitter(height = 2, width = 0.1, seed = 233),
                 dotsize=0.7, aes(fill = Hotspots), color = 'black',stroke=0.8) + 
    scale_fill_manual(values= c("#E84D3F","#77BFD1","#DEA15B"))+
    scale_color_manual(values= c("#E84D3F","#77BFD1","#DEA15B"))+
    ylim(c(0,100)) + ylab(paste0('% of ', cell_name)) +
    theme_bw() + 
    theme(axis.text.x = element_blank(), legend.position = 'None', axis.title=element_text(size=18),
          axis.text=element_text(size=16, colour="black"),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.background = element_blank(), aspect.ratio = 1.6/1.3,
          axis.ticks.length=unit(.25, "cm")) + xlab('') +
    annotate(geom = 'segment', x = 1, xend = 2, y = 60, yend = 60, color = subset(compare_hs, variable == cell_type)$line_color[1])+
    geom_text(x = 1.5, y = 63, label = subset(compare_hs, variable == cell_type)$p_adj_star[1]) +
    annotate(geom = 'segment', x = 1, xend = 3, y = 75, yend = 75, color = subset(compare_hs, variable == cell_type)$line_color[2])+
    geom_text(x = 2, y = 78, label = subset(compare_hs, variable == cell_type)$p_adj_star[2]) +
    annotate(geom = 'segment', x = 2, xend = 3, y = 90, yend = 90, color = subset(compare_hs, variable == cell_type)$line_color[3])+
    geom_text(x = 2.5, y = 93, label = subset(compare_hs, variable == cell_type)$p_adj_star[3])
  return(p)
}

plt_region_celltype <- function(df, compare_region, cell_type = 'cd8', cell_name = 'CD8+ T cells'){
  p <- ggplot(subset(df,Hotspots %in% c('Immune') & variable == cell_type), aes(x= Region, y = value)) +
    geom_boxplot(outlier.shape = NA, aes(fill = Region))+
    geom_dotplot(binaxis='y', stackdir='center', stackratio = 0.3,
                 position=position_jitter(height = 2, width = 0.1, seed = 233),
                 dotsize=0.7, aes(fill = Region), color = 'black',stroke=0.8) + 
    scale_fill_brewer(palette = "Set2")+
    scale_color_brewer(palette = "Set2")+
    ylim(c(0,100)) + ylab(paste0('% of ', cell_name)) +
    theme_bw() + 
    theme(axis.text.x = element_blank(), legend.position = 'None', axis.title=element_text(size=18),
          axis.text=element_text(size=16, colour="black"),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
          axis.line = element_line(colour = "black"),
          panel.background = element_blank(), aspect.ratio = 1.6/1.3,
          axis.ticks.length=unit(.25, "cm")) + xlab('') +
    annotate(geom = 'segment', x = 1, xend = 2, y = 70, yend = 70, color = subset(compare_region, variable == cell_type)$line_color[1])+
    geom_text(x = 1.5, y = 73, label = subset(compare_region, variable == cell_type)$p_adj_star[1]) +
    annotate(geom = 'segment', x = 1, xend = 3, y = 85, yend = 85, color = subset(compare_region, variable == cell_type)$line_color[2])+
    geom_text(x = 2, y = 88, label = subset(compare_region, variable == cell_type)$p_adj_star[2]) +
    annotate(geom = 'segment', x = 2, xend = 3, y = 100, yend = 100, color = subset(compare_region, variable == cell_type)$line_color[3])+
    geom_text(x = 2.5, y = 103, label = subset(compare_region, variable == cell_type)$p_adj_star[3])
  
  return(p)
}

plt_sur_by_smokingClass <- function(LUSCmean, class_index, title){
  
  df <- subset(LUSCmean, smokingClass == class_index)
  cut <- surv_cutpoint(df, time = "time", event = "event",
                       variables = c("fi"))
  df_cut <- surv_categorize(cut)
  
  fit <- survfit(Surv(time, event)~fi, data = df_cut)
  p<-ggsurvplot(fit, data = df_cut,
                ylab = "Overall Survival",
                xlab = "Time (Days)",
                break.time.by = 500, xlim = c(0,3000),
                conf.int = F,
                risk.table = T, risk.table.y.text.col = T, risk.table.y.text = F, risk.table.height = 0.3, 
                risk.table.title = "No. at Risk", risk.table.fontsize = 4, 
                pval = T, pval.coord = c(0.1, 0.1,0.1,0.1),
                legend = c(0.85, 0.8), title = title)
  return(p) 
}

cus_theme <-theme(plot.title=element_text(size=16),
                  axis.text=element_text(size=14),
                  axis.title=element_text(size=16),
                  legend.title = element_text(size = 14),
                  legend.text = element_text(size = 14))

```

#Table 1. Characteristics of the included patients and their association with the Sintra/immune, Sintra/cancer, and Sintra/tissue.
```{r}
clin <- survivalTCGA(LUAD.clinical, LUSC.clinical, 
                     extract.cols=c("admin.disease_code", "patient.days_to_death", "patient.stage_event.pathologic_stage", "patient.histological_type",
                                    "patient.tobacco_smoking_history", "patient.number_pack_years_smoked", "patient.year_of_tobacco_smoking_onset",
                                    "patient.stopped_smoking_year", "patient.residual_tumor", "patient.primary_therapy_outcome_success",
                                    "patient.age_at_initial_pathologic_diagnosis", "patient.drugs.drug.therapy_types.therapy_type", "patient.gender"))

df_LUAD_LUSC <- rbind(LUADmean, LUSCmean)

df_LUAD_LUSC <- merge(df_LUAD_LUSC, clin[,c('bcr_patient_barcode', 'patient.gender')], all.x = T)


#=======Histology type=======#
# df_LUAD_LUSC$fi_group <- ifelse(df_LUAD_LUSC$fi > 0.326, "High", "Low" )
df_LUAD_LUSC$fi_group <- ifelse(df_LUAD_LUSC$fi > median(df_LUAD_LUSC$fi, na.rm = T), "High", "Low" )
df_LUAD_LUSC$fc_group <- ifelse(df_LUAD_LUSC$fc > median(df_LUAD_LUSC$fc, na.rm = T), "High", "Low" )
df_LUAD_LUSC$ft_group <- ifelse(df_LUAD_LUSC$ft > median(df_LUAD_LUSC$ft, na.rm = T), "High", "Low" )

s = nrow(subset(df_LUAD_LUSC, !is.na(admin.disease_code)))
table(df_LUAD_LUSC$admin.disease_code)
round(table(df_LUAD_LUSC$admin.disease_code) / s * 100, 2)

table(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')$fi_group)
round(table(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')$fi_group) / nrow(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')) * 100, 2)

table(subset(df_LUAD_LUSC, admin.disease_code == 'luad')$fi_group)
round(table(subset(df_LUAD_LUSC, admin.disease_code == 'luad')$fi_group) / nrow(subset(df_LUAD_LUSC, admin.disease_code == 'luad')) * 100, 2)

table(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')$fc_group)
round(table(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')$fi_group) / nrow(subset(df_LUAD_LUSC, admin.disease_code == 'lusc')) * 100, 2)

table(subset(df_LUAD_LUSC, admin.disease_code == 'luad')$fc_group)
round(table(subset(df_LUAD_LUSC, admin.disease_code == 'luad')$fi_group) / nrow(subset(df_LUAD_LUSC, admin.disease_code == 'luad')) * 100, 2)

ggboxplot(df_LUAD_LUSC, x = "admin.disease_code", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(df_LUAD_LUSC, x = "admin.disease_code", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(df_LUAD_LUSC, x = "admin.disease_code", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

#=======Age=======#
df_LUAD_LUSC$age_group <- ifelse(df_LUAD_LUSC$age >=65, ">=65", "<65")
s = nrow(subset(df_LUAD_LUSC, !is.na(age_group)))
table(df_LUAD_LUSC$age_group)
round(table(df_LUAD_LUSC$age_group) / nrow(subset(df_LUAD_LUSC, !is.na(age_group))) * 100, 2)

table(subset(df_LUAD_LUSC, age_group == '>=65')$fi_group)
round(table(subset(df_LUAD_LUSC, age_group == '>=65')$fi_group) / nrow(subset(df_LUAD_LUSC, age_group == '>=65')) * 100, 2)

table(subset(df_LUAD_LUSC, age_group == '<65')$fi_group)
round(table(subset(df_LUAD_LUSC, age_group == '<65')$fi_group) / nrow(subset(df_LUAD_LUSC, age_group == '<65')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(age_group)), x = "age_group", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(age_group)), x = "age_group", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(age_group)), x = "age_group", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

#=======sex=======#
s = nrow(subset(df_LUAD_LUSC, !is.na(patient.gender)))
table(df_LUAD_LUSC$patient.gender)
round(table(df_LUAD_LUSC$patient.gender) / nrow(subset(df_LUAD_LUSC, !is.na(patient.gender))) * 100, 2)

table(subset(df_LUAD_LUSC, patient.gender == 'male')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.gender == 'male')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.gender == 'male')) * 100, 2)

table(subset(df_LUAD_LUSC, patient.gender == 'female')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.gender == 'female')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.gender == 'female')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.gender)), x = "patient.gender", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.gender)), x = "patient.gender", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.gender)), x = "patient.gender", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

#=======Smoking history=======#
s = nrow(subset(df_LUAD_LUSC, !is.na(patient.tobacco_smoking_history)))
round(table(subset(df_LUAD_LUSC, !is.na(patient.tobacco_smoking_history) & patient.tobacco_smoking_history != "current reformed smoker, duration not specified")$patient.tobacco_smoking_history) / (s-9) * 100, 2)

table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for < or = 15 years')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for < or = 15 years')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for < or = 15 years')) * 100, 2)

table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for > 15 years')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for > 15 years')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current reformed smoker for > 15 years')) * 100, 2)

table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current smoker')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current smoker')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'current smoker')) * 100, 2)

table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'lifelong non-smoker')$fi_group)
round(table(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'lifelong non-smoker')$fi_group) / nrow(subset(df_LUAD_LUSC, patient.tobacco_smoking_history == 'lifelong non-smoker')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.tobacco_smoking_history)), x = "patient.tobacco_smoking_history", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means()

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.tobacco_smoking_history)), x = "patient.tobacco_smoking_history", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means()

ggboxplot(subset(df_LUAD_LUSC, !is.na(patient.tobacco_smoking_history)), x = "patient.tobacco_smoking_history", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means()

#=======pack years=======#
s = nrow(subset(df_LUAD_LUSC, !is.na(pack_years)))
df_LUAD_LUSC$pack_year_group <- ifelse(df_LUAD_LUSC$pack_years >=35, ">=35", "<35")
table(df_LUAD_LUSC$pack_year_group)
round(table(df_LUAD_LUSC$pack_year_group) / s * 100, 2)

table(subset(df_LUAD_LUSC, pack_year_group == '>=35')$fi_group)
round(table(subset(df_LUAD_LUSC, pack_year_group == '>=35')$fi_group) / nrow(subset(df_LUAD_LUSC, pack_year_group == '>=35')) * 100, 2)

table(subset(df_LUAD_LUSC, pack_year_group == '<35')$fi_group)
round(table(subset(df_LUAD_LUSC, pack_year_group == '<35')$fi_group) / nrow(subset(df_LUAD_LUSC, pack_year_group == '<35')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(pack_year_group)), x = "pack_year_group", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(pack_year_group)), x = "pack_year_group", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(pack_year_group)), x = "pack_year_group", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

#=======TMB=======#
s = nrow(subset(df_LUAD_LUSC, !is.na(total_perMB)))
df_LUAD_LUSC$total_perMB_group <- ifelse(df_LUAD_LUSC$total_perMB >=3, ">=3", "<3")
table(df_LUAD_LUSC$total_perMB_group)
round(table(df_LUAD_LUSC$total_perMB_group) / s * 100, 2)

table(subset(df_LUAD_LUSC, total_perMB_group == '>=3')$fi_group)
round(table(subset(df_LUAD_LUSC, total_perMB_group == '>=3')$fi_group) / nrow(subset(df_LUAD_LUSC, total_perMB_group == '>=3')) * 100, 2)

table(subset(df_LUAD_LUSC, total_perMB_group == '<3')$fi_group)
round(table(subset(df_LUAD_LUSC, total_perMB_group == '<3')$fi_group) / nrow(subset(df_LUAD_LUSC, total_perMB_group == '<3')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(total_perMB_group)), x = "total_perMB_group", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(total_perMB_group)), x = "total_perMB_group", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

ggboxplot(subset(df_LUAD_LUSC, !is.na(total_perMB_group)), x = "total_perMB_group", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means(method = "wilcox")

#=======stages=======#
s = nrow(subset(df_LUAD_LUSC, !is.na(stage_merge)))
table(df_LUAD_LUSC$stage_merge)
round(table(df_LUAD_LUSC$stage_merge) / s * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'ia')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'ia')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'ia')) * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'ib')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'ib')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'ib')) * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'iia')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'iia')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'iia')) * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'iib')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'iib')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'iib')) * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'iiia')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'iiia')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'iiia')) * 100, 2)

table(subset(df_LUAD_LUSC, stage_merge == 'iv')$fi_group)
round(table(subset(df_LUAD_LUSC, stage_merge == 'iv')$fi_group) / nrow(subset(df_LUAD_LUSC, stage_merge == 'iv')) * 100, 2)

ggboxplot(subset(df_LUAD_LUSC, !is.na(stage_merge)), x = "stage_merge", y = "fi",
               add = "jitter", border = "white")+
  stat_compare_means()

ggboxplot(subset(df_LUAD_LUSC, !is.na(stage_merge)), x = "stage_merge", y = "fc",
               add = "jitter", border = "white")+
  stat_compare_means()

ggboxplot(subset(df_LUAD_LUSC, !is.na(stage_merge)), x = "stage_merge", y = "ft",
               add = "jitter", border = "white")+
  stat_compare_means()
```

# Fig. 1  Immune hotspots in LUSC
## Fig. 1C
```{r}
#Optimize threshold for the discovery set
LUSCdisc <- subset(LUSCmean, cohort == "disc")
LUSCvalid <- subset(LUSCmean, cohort == "valid")

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("fi", "lym_per"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "fiC", "fi", res.cutLUSCd)
LUSCvalid <- stratify_df(LUSCvalid, "lym_perC", "lym_per", res.cutLUSCd)

#survival plot
fit <- survfit(Surv(time, event)~fiC, data = LUSCvalid)

ggsurvplot(fit, data = LUSCvalid, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
```

## Fig. 1D
```{r}
res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("fi", "lym_per"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "fiC", "fi", res.cutLUSCd)
LUSCvalid <- stratify_df(LUSCvalid, "lym_perC", "lym_per", res.cutLUSCd)

fitd <- coxph(Surv(time, event)~fiC+age+stage_merge+pack_years, data = LUSCvalid)
ggforest(fitd, main = "TCGA LUSC: validation cohort")

```
## Fig. 1E
```{r}
ggscatter(LUSCmean, 
          x = "fi", y = "lym_per", 
          add = "reg.line", 
          xlab = expression('S'['intra/immune']), 
          ylab = "Lymphocyte%",
          color = "blue4",
          conf.int = TRUE, size = 2,
          add.params = list(color = "grey50", fill = "azure3"), 
          cor.coef = TRUE, cor.method = "spearman") + cus_theme
```

## Fig. 1F
```{r}

res.cutLUSCd <- surv_cutpoint(LUSCmean, time = "time", event = "event",minprop = 0.2,
                              variables = c("fi", "lym_per"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCmean <- stratify_df(LUSCmean, "fiC", "fi", res.cutLUSCd)
LUSCmean <- stratify_df(LUSCmean, "lym_perC", "lym_per", res.cutLUSCd)

LUSCmean$fi_lym <- ifelse(LUSCmean$lym_perC == "High" & LUSCmean$fiC == "Low", "Low score high lym%", "All others")
LUSCmean$fi_lym <- factor(LUSCmean$fi_lym, levels = c("Low score high lym%", "All others"))

fit <- survfit(Surv(time, event)~fi_lym, data = LUSCmean)

ggsurvplot(fit, data = LUSCmean, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))


LUSCmean$fi_lym <- factor(LUSCmean$fi_lym, levels = c("All others", "Low score high lym%"))
fitd <- coxph(Surv(time, event)~fi_lym, data = LUSCmean)
ggforest(fitd, main = "TCGA LUSC: validation cohort")

```

#Fig.2
## Fig. 2A
```{r}
# Enrichment Analysis EA (TCGAVisualize)
# Gene Ontology (GO) and Pathway enrichment barPlot
TCGAvisualize_EAbarplot(tf = rownames(ansEA$ResBP), 
                        GOBPTab = ansEA$ResBP,
                        GOCCTab = ansEA$ResCC,
                        GOMFTab = ansEA$ResMF,
                        PathTab = ansEA$ResPat,
                        nRGTab = Genelist, 
                        nBar = 10,
                        filename=file.path(save_dir, "TCGAvisualize_EAbarplot_fiLUSC_updated.pdf"))
```

## Fig. 2B
```{r}
TCGAVisualize_volcano(x = dataDEGs$logFC,
                      y = dataDEGs$FDR,
                      x.cut = 3,
                      y.cut = 10^-5,
                      names = rownames(dataDEGs),
                      color = c("black","red","darkgreen"),
                      names.size = 2,
                      xlab = " Gene expression fold change (Log2)",
                      legend = "State",
                      highlight = c("CD19", "CD79B", "MS4A1", "CD20", "CXCR5", "FCER2"),
                      title = "Volcano plot (fi-high vs fi-low)",
                      show.names = "both",
                      width = 10,
                      filename = file.path(save_dir, "volcano_fiLUSC_verySigBcell_FCER2_remove_xth.pdf"))

```

## Fig. 2C
```{r}
LUSCmean$fiC_median <- ifelse(LUSCmean$fi >= median(LUSCmean$fi, na.rm = T), "High", "Low")

LUSCmean_B_cell <- merge(LUSCmean_B_cell, LUSCmean[,c("bcr_patient_barcode", "fiC_median")], all.x = T)

df_m <- melt(LUSCmean_B_cell)

#P value adjustment
compare_means(value ~ fiC_median, data = df_m, group.by = "variable", p.adjust.method = "BH")

ggboxplot(df_m, x = "fiC_median", y = "value", facet.by = "variable",#title = "EPIC",
             xlab = "S_intra/immune",size=0.5, scales = "free_y",
             color = "fiC_median", palette = c("#EE0000", "#00BFFF"), 
             add = "jitter", border = "white")+
            stat_compare_means(label = "p.format", label.x = 1.3)+
            theme(text = element_text(size=18))+
            theme(legend.position="")+
            theme(plot.title = element_text(hjust = 0.5))


```
#Fig. 4
## Fig. 4B
```{r}
df <- subset(sum_valid, Hotspots %in% c("Cancer-immune", "Immune"))

df <- df %>%
  group_by(slide, Hotspots) %>%
  summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
            foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE),
            cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE), area = paste(x.s, y.s) %>% unique %>% length())

#devide by the total area of hotspots

df[,setdiff(colnames(df), c('slide','Hotspots', 'area'))] <- df[,setdiff(colnames(df), c('slide','Hotspots', 'area'))] / df$area / (50*50) * 10^6 
df$area <- NULL

df <- melt(df, id = c('slide','Hotspots'))

df$Hotspots <- factor(df$Hotspots , levels = c('Cancer-immune','Immune'))

#statistic
compare_hs <- compare_means(value ~ Hotspots, p.adjust.method = "BH", method='wilcox.test', paired = T,
                            group.by = "variable",
                            data =  df) %>%
  mutate(y_pos = rep(75,6), p.adj = format.pval(p.adj, digits = 2),
         p_adj_star = ifelse(as.numeric(gsub("< ", "", p.adj)) <0.0001, '****', 
                             ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.001, '***', 
                                    ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.01, '**', 
                                           ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.05, '*', '')))))  %>%
  mutate(line_color = ifelse(p_adj_star == "", "white", "black"))

compare_hs

ggplot(df, aes(x= Hotspots, y = value)) + facet_grid(. ~ variable) + 
  geom_boxplot(outlier.shape = NA, aes(fill = Hotspots))+
  geom_dotplot(binaxis='y', stackdir='center', stackratio = 0.8,
               position=position_jitter(height = 2, width = 0.1, seed = 233),
               dotsize=1, aes(fill = Hotspots), color = 'black',stroke=0.4) + 
  scale_fill_manual(values= c("#A0C588","#636AA8"))+
  scale_color_manual(values= c("#A0C588","#636AA8"))+
  ylab(bquote(.(paste0("Cell counts", " /")) ~ mm^2)) +
  theme_bw() + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank() ,
        legend.position = 'None', axis.title=element_text(size=18),
        axis.text=element_text(size=16, colour="black"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(), aspect.ratio = 1.6/1.3,
        axis.ticks.length=unit(.25, "cm")) + xlab('')
```

## Fig. 4C
```{r}
#sum by region and get density
df <- subset(sum_valid, Hotspots %in% c("Cancer-immune", "Immune"))

df <- df %>%
  group_by(slide, Region) %>%
  summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
            foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE),
            cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE), area = paste(x.s, y.s) %>% unique %>% length())

df[,setdiff(colnames(df), c('slide','area', 'Region'))] <- df[,setdiff(colnames(df), c('slide','area', 'Region'))] / df$area / (50*50) * 10^6 
df$area <- NULL

df <- melt(df, id = c('slide','Region'))

df$Region <- factor(df$Region  , levels = c('TLS','LAG', 'UD'))

#statistic

df <- tidyr::complete(df, slide, Region, variable, fill = list(value=NA))

compare_region <- compare_means(value ~ Region, p.adjust.method = "BH", method='wilcox.test', paired = T,
                                group.by = "variable",
                                data =  df) %>%
  mutate(y_pos =rep(c(80, 95, 110),6), p.adj = format.pval(p.adj, digits = 2),
         p_adj_star = ifelse(as.numeric(gsub("< ", "", p.adj)) <0.0001, '****', 
                             ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.001, '***', 
                                    ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.01, '**', 
                                           ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.05, '*', ''))))) %>%
  mutate(line_color = ifelse(p_adj_star == "", "white", "black"))

compare_region

ggplot(df, aes(x= Region, y = value)) + facet_grid(. ~ variable) + 
  geom_boxplot(outlier.shape = NA, aes(fill = Region))+
  geom_dotplot(binaxis='y', stackdir='center', stackratio = 0.8,
               position=position_jitter(height = 2, width = 0.1, seed = 233),
               dotsize=1, aes(fill = Region), color = 'black',stroke=0.4) + 
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Set2")+
  ylab(bquote(.(paste0("Cell counts", " /")) ~ mm^2)) +
  
  theme_bw() + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank() ,
        legend.position = 'None', axis.title=element_text(size=18),
        axis.text=element_text(size=16, colour="black"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.background = element_blank(), aspect.ratio = 1.6/1.3,
        axis.ticks.length=unit(.25, "cm")) + xlab('')

```

## Fig. 4D
```{r}
df_region_ratio <- sum_valid %>% group_by(Hotspots, Region, slide) %>% summarise(n = n())

df_region_ratio <- subset(df_region_ratio, Hotspots %in% c('Cancer-immune','Immune'))

df_region_ratio_sum <- df_region_ratio %>% group_by(slide, Hotspots) %>% summarise(sum_n = sum(n))

df_region_ratio <- merge(df_region_ratio, df_region_ratio_sum, by = c('slide', 'Hotspots'))

df_all <- df_region_ratio %>% group_by(Region, Hotspots) %>% summarise(n_region = sum(n, na.rm = T))
df_all$Region <- factor(df_all$Region, levels = c("TLS", "LAG", "UD"))

#Peritumoral IH
df <- subset(df_all, Hotspots == 'Immune') 

df$labs <- paste0(df$n_region, "/", sum(df$n_region))
df$labs <- df$n_region / sum(df$n_region) * 100
df$labs <- paste0(round(df$labs,2), '%')
df <- as.data.frame(df)
ggpie(df, "n_region", label = 'labs', lab.pos = "out", lab.font = "white",
      fill = "Region", color = "Gray",
      palette = 'Set2', title = 'Composition of Regions at peritumoral immune hotspots')

#Intratumoral IH
df <- subset(df_all, Hotspots == 'Cancer-immune') 

df$labs <- paste0(df$n_region, "/", sum(df$n_region))
df$labs <- df$n_region / sum(df$n_region) * 100
df$labs <- paste0(round(df$labs,2), '%')
df <- as.data.frame(df)
ggpie(df, "n_region", label = 'labs', lab.pos = "out", lab.font = "white",
      fill = "Region", color = "Gray",
      palette = 'Set2', title = 'Composition of Regions at intratumoral immune hotspots')

```

## Fig. 4E
```{r}
df_region_ratio <- sum_valid %>% group_by(Hotspots, Region, slide) %>% summarise(n = n())

df_region_ratio <- subset(df_region_ratio, Hotspots %in% c('Cancer-immune','Immune'))

df_region_ratio_sum <- df_region_ratio %>% group_by(slide, Hotspots) %>% summarise(sum_n = sum(n))
df_region_ratio <- merge(df_region_ratio, df_region_ratio_sum, by = c('slide', 'Hotspots'))
df_region_ratio$ratio <- df_region_ratio$n / df_region_ratio$sum_n * 100

df_region_ratio <- tidyr::complete(df_region_ratio, slide, Region, Hotspots, fill = list(ratio=0))

#TLS
#average in peritumoral IH
mean(subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Immune'))$ratio)
#sd in peritumoral IH
sd(subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Immune'))$ratio)

#average in intratumoral IH
mean(subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Cancer-immune'))$ratio)
#sd in intratumoral IH
sd(subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Cancer-immune'))$ratio)

df_p <- compare_means(ratio ~ Hotspots, p.adjust.method = "BH", method='wilcox.test', paired = T, 
                      data = subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Cancer-immune','Immune'))) 

p <- ggboxplot(subset(df_region_ratio, Region == 'TLS' & Hotspots %in% c('Cancer-immune','Immune')), x = 'Hotspots', y = 'ratio', 
          fill = 'Hotspots', palette = c("#EE0000","#00BFFF"), add = "jitter", add.params = list(color="black", size=1)) +
  theme_bw() + ylab('% of area') +
  theme( axis.title=element_text(size=18),
         axis.text=element_text(size=16, colour="black"),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
         axis.line = element_line(colour = "black"),
         panel.background = element_blank(), aspect.ratio = 1.6/1.3, 
         axis.ticks.length=unit(.25, "cm"), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  geom_text(x=1.5, y = 15, label=paste0("p = ", df_p$p.adj), color = 'black', size=7)


#LAG

#average in peritumoral IH
mean(subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Immune'))$ratio)
#sd in peritumoral IH
sd(subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Immune'))$ratio)

#average in intratumoral IH
mean(subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Cancer-immune'))$ratio)
#sd in intratumoral IH
sd(subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Cancer-immune'))$ratio)

df_p <- compare_means(ratio ~ Hotspots, p.adjust.method = "BH", method='wilcox.test', paired = T, 
                      data = subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Cancer-immune','Immune'))) 

p <- ggboxplot(subset(df_region_ratio, Region == 'LAG' & Hotspots %in% c('Cancer-immune','Immune')), x = 'Hotspots', y = 'ratio', 
          fill = 'Hotspots', palette = c("#EE0000","#00BFFF"), add = "jitter", add.params = list(color="black", size=1)) +
  theme_bw() + ylab('% of area') +
  theme( axis.title=element_text(size=18),
         axis.text=element_text(size=16, colour="black"),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
         axis.line = element_line(colour = "black"),
         panel.background = element_blank(), aspect.ratio = 1.6/1.3, 
         axis.ticks.length=unit(.25, "cm"), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  geom_text(x=1.5, y = 8.5, label=paste0("p = ", df_p$p.adj), color = 'black', size=7)

```
# Fig. 5
## Fig. 5B
```{r}
library(vegan)
df_inter_den <- df_inter_prop
df_den <- df_inter_den[,which(colnames(df_inter_den) == "cd20") : which(colnames(df_inter_den) == "foxp3")]
df_inter <- df_inter_den[,which(colnames(df_inter_den) == "cd20cxcr5.cd4") : which(colnames(df_inter_den) == "cd20.cd79bCoexp")]

df_den$shannon_den <- diversity(df_den[,1:6], index = "shannon")
df_inter$shannon_inter <- diversity(df_inter, index = "shannon")
df_den <- cbind(df_den, df_inter_den[,c("hs", "clump_id", "slide", "clump_area", "hs_dig")])
df_inter <- cbind(df_inter, df_inter_den[,c("hs", "clump_id", "slide", "clump_area", "hs_dig")])

df_p <- compare_means(shannon_den ~ hs_dig, p.adjust.method = "BH", method='wilcox.test', paired = F,
                      data = df_den)

df_p_inter <- compare_means(shannon_inter ~ hs_dig, p.adjust.method = "BH", method='wilcox.test', paired = F, 
                      data = df_inter)

df_den$shannon_inter <- df_inter$shannon_inter

colnames(df_den)[which(colnames(df_den) == 'hs')] <- "Hotspots"
levels(df_den$Hotspots) <- c("Intra", "Peri")

df <- df_den[,c('shannon_den', 'shannon_inter', 'Hotspots')]

df <- melt(df)

p <- ggboxplot(df, x = 'Hotspots', y = 'value', color = 'Hotspots', palette = c("#A0C588","#636AA8"), add = "jitter", facet.by = "variable") +
  theme_bw() + ylab('Shannon diversity') +
  theme( axis.title=element_text(size=18),
         axis.text=element_text(size=16, colour="black"),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.border = element_blank(),
         axis.line = element_line(colour = "black"),
         panel.background = element_blank(), aspect.ratio = 1.6/1.3, 
         axis.ticks.length=unit(.25, "cm"), axis.text.x = element_blank(), axis.title.x = element_blank(),
         axis.ticks.x = element_blank()) +
  geom_text(x=1.5, y = 2, label=paste0("p = ", df_p_inter$p.adj), color = 'black', size=7) +
  ylim(c(-0.1,2))

```

## Fig. 5C
```{r}
df_inter <- df_inter_den[, which(colnames(df_inter_den) == "cd20cxcr5.cd4") : which(colnames(df_inter_den) == "cd20.cd79bCoexp")]
variable_list <- colnames(df_inter)

df_inter$hs_dig <- df_inter_den$hs_dig

coef_all <- data.frame()
df <- df_inter

for (v in variable_list){
  print(v)
  logit  <- glm(as.formula(paste("hs_dig~", v)), data = df, family = "binomial")
  
  sum_df <- summary(logit)
  
  sum_df_coef <- data.frame(sum_df$coefficients)
  
  coef_all <- rbind(coef_all, sum_df_coef[2,])
}

coef_all$p_adj <- p.adjust(coef_all$Pr...z.., method = "BH")

coef_all <- coef_all %>% mutate(p_star = ifelse(as.numeric(gsub("< ", "", Pr...z..)) <0.0001, '****', 
                                                ifelse(as.numeric(gsub("< ", "", Pr...z..)) < 0.001, '***', 
                                                       ifelse(as.numeric(gsub("< ", "", Pr...z..)) < 0.01, '**', 
                                                              ifelse(as.numeric(gsub("< ", "", Pr...z..)) < 0.05, '*', ''))))) %>%
  mutate(p_adj_star = ifelse(as.numeric(gsub("< ", "", p_adj)) <0.0001, '****', 
                             ifelse(as.numeric(gsub("< ", "", p_adj)) < 0.001, '***', 
                                    ifelse(as.numeric(gsub("< ", "", p_adj)) < 0.01, '**', 
                                           ifelse(as.numeric(gsub("< ", "", p_adj)) < 0.05, '*', ''))))) 


coef_all$Variable <- rownames(coef_all)
coef_all$color <- ifelse(coef_all$Estimate<0 & coef_all$p_adj_star != "", "red", "grey")
coef_all$log_q <- -log10(coef_all$p_adj)

p <- ggscatter(data = coef_all, x = "Variable", y = "Estimate", color = "color", size = 4, palette = c("grey70", "maroon")) +coord_flip() + theme(axis.text.x = element_text(angle = 90))+
  geom_hline(yintercept = 0, linetype = 2) + ylim(c(-30, 30))

```

# Fig. 6
```{r}
df_cell_cd8_foxp3_bcell <- sum_valid

df_cell_cd8_foxp3_bcell$sum_cells <- rowSums(df_cell_cd8_foxp3_bcell[,c(which(colnames(df_cell_cd8_foxp3_bcell)=='cd4'):which(colnames(df_cell_cd8_foxp3_bcell)=='foxp3'),which(colnames(df_cell_cd8_foxp3_bcell)=='cd20'):which(colnames(df_cell_cd8_foxp3_bcell)=='cd79bCoexp'))])

df_cell_cd8_foxp3_bcell$bcells <- rowSums(df_cell_cd8_foxp3_bcell[,which(colnames(df_cell_cd8_foxp3_bcell)=='cd20'):which(colnames(df_cell_cd8_foxp3_bcell)=='cd79bCoexp')])

df_cell_cd8_foxp3_bcell_cl <- subset(df_cell_cd8_foxp3_bcell, Hotspots == "Cancer-immune")
df_cell_cd8_foxp3_bcell_l <- subset(df_cell_cd8_foxp3_bcell, Hotspots == "Immune")

df <- df_cell_cd8_foxp3_bcell_l %>% group_by(slide) %>% summarise(n = n())

df_cell_cd8_foxp3_bcell_cl$Hotspots <- 'Intratumoral IH'
df_cell_cd8_foxp3_bcell_l$Hotspots <- 'Peritumoral IH'


select_col <- c('slide', 'Hotspots', 'Region', 'clump_id', 'cd8', 'cd4', 'foxp3', 'cd20', 'cd20cxcr5', 'cd79bCoexp', 'sum_cells', 'bcells')

df_cl_l <- rbind(df_cell_cd8_foxp3_bcell_l[, select_col], df_cell_cd8_foxp3_bcell_cl[, select_col])

#exclude TLS and LAG
df_cl_l <- subset(df_cl_l, Region == "UD")

df_cl_l <- df_cl_l %>% group_by(slide, Hotspots, clump_id) %>% summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE), foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE), cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE),sum_cells = sum(sum_cells, na.rm = TRUE))


df_cl_l$bcells <- rowSums(df_cl_l[,c('cd20', 'cd20cxcr5', 'cd79bCoexp')]) / df_cl_l$sum_cells

df_cl_l[,c('cd8', 'cd4', 'foxp3', 'cd20', 'cd20cxcr5', 'cd79bCoexp')] <- df_cl_l[,c('cd8', 'cd4', 'foxp3', 'cd20', 'cd20cxcr5', 'cd79bCoexp')] / df_cl_l$sum_cells

#======CD8/FOXP3 between intra and peritumoral======#
df_cl_l$cd8_foxp3 <-  as.numeric(df_cl_l$cd8 / df_cl_l$foxp3)

df <- subset(df_cl_l, Hotspots == "Intratumoral IH")
df <- subset(df, is.finite(cd8_foxp3))
sd(df$cd8_foxp3, na.rm = T)
mean(df$cd8_foxp3, na.rm = T)

df <- subset(df_cl_l, Hotspots == "Peritumoral IH")
df <- subset(df, is.finite(cd8_foxp3))
sd(df$cd8_foxp3, na.rm = T)
mean(df$cd8_foxp3, na.rm = T)

p <- ggboxplot(subset(df_cl_l, !is.na(cd8_foxp3)), y = 'cd8_foxp3', x = 'Hotspots', add = 'jitter', color = 'Hotspots', palette = c("firebrick4", "tan3")) + 
  stat_compare_means(label = "p.format", label.x = 1.4, paired = F, label.y = 50) + 
  xlab('') +
  cus_theme


df_cl_l_1 <- subset(df_cl_l, Hotspots == 'Intratumoral IH')
df_cl_l_2 <- subset(df_cl_l, Hotspots == 'Peritumoral IH')

#assign high/low for intratumoral and peritumoral IH respectively
df_cl_l_1$cd8_foxp3_class <- ifelse(df_cl_l_1$cd8_foxp3 >= median(df_cl_l_1$cd8_foxp3, na.rm = T), "High", "Low")
df_cl_l_2$cd8_foxp3_class <- ifelse(df_cl_l_2$cd8_foxp3 >= median(df_cl_l_2$cd8_foxp3, na.rm = T), "High", "Low")

df_cl_l_1$foxp3_class <- ifelse(df_cl_l_1$foxp3 >= median(df_cl_l_1$foxp3, na.rm = T), "High", "Low")
df_cl_l_2$foxp3_class <- ifelse(df_cl_l_2$foxp3 >= median(df_cl_l_2$foxp3, na.rm = T), "High", "Low")

df_cl_l_1$cd8_class <- ifelse(df_cl_l_1$cd8 >= median(df_cl_l_1$cd8, na.rm = T), "High", "Low")
df_cl_l_2$cd8_class <- ifelse(df_cl_l_2$cd8 >= median(df_cl_l_2$cd8, na.rm = T), "High", "Low")

df_cl_l_1$bcell_class <- ifelse(df_cl_l_1$bcells >= median(df_cl_l_1$bcells, na.rm = T), "High", "Low")
df_cl_l_2$bcell_class <- ifelse(df_cl_l_2$bcells >= median(df_cl_l_2$bcells, na.rm = T), "High", "Low")

df_cl_l_3 <- rbind(df_cl_l_1, df_cl_l_2)

df_cl_l_3$cd8_foxp3_class <- factor(df_cl_l_3$cd8_foxp3_class, levels = c("High", "Low"))
df_cl_l_3$foxp3_class <- factor(df_cl_l_3$foxp3_class, levels = c("High", "Low"))
df_cl_l_3$cd8_class <- factor(df_cl_l_3$cd8_class, levels = c("High", "Low"))
df_cl_l_3$bcell_class <- factor(df_cl_l_3$bcell_class, levels = c("High", "Low"))
```

## Fig. 6A
```{r}

#cd8/foxp3 high low
cell_class <- c("cd20", "cd20cxcr5", "cd79bCoexp")

cell_label <- c("% of CD20+CXCR5- B cells", "% of CD20+CXCR5+ B cells", "% of CD79b+ B cells")
# cell_label <- c("% of CD20+CXCR5+ B cells")

for (i in 1:length(cell_class)){
  print(cell_class[i])
  pdf(file.path(save_dir, paste0(cell_class[i], "_cd8_to_foxp3_class.pdf")))
  
  p <- ggboxplot(subset(df_cl_l_3, !is.na(cd8_foxp3_class)), x = 'cd8_foxp3_class', y = cell_class[i], add = 'jitter', color = 'cd8_foxp3_class', palette = c("firebrick4", "tan3"), facet.by = "Hotspots") + 
    stat_compare_means(label = "p.format", label.x = 1.3, paired = F) + #label.y = 0.35
    xlab('') + ylab(cell_label[i]) #+ ylim(c(0,0.4))
    cus_theme
  
  print(p)
  dev.off()
  
}

```

## Fig. 6B
```{r}
#bcell perc high low
cell_class <- c("foxp3", "cd8")
cell_label <- c("% of CD4+FOXP3+ T cells","% of CD8+ T cells")

for (i in 1:length(cell_class)){
  print(cell_class[i])
  pdf(file.path(save_dir, paste0(cell_class[i], "_bcell_class.pdf")))
  
  p <- ggboxplot(subset(df_cl_l_3, !is.na(bcell_class)), x = 'bcell_class', y = cell_class[i], add = 'jitter', color = 'bcell_class', palette = c("firebrick4", "tan3"), facet.by = "Hotspots") + 
    stat_compare_means(label = "p.format", label.x = 1.3, paired = F) + 
    xlab('') + ylab(cell_label[i]) +
    cus_theme
  
  print(p)
  dev.off() 
  
}
```

## Fig. 6C
```{r}
cell_class <- c("cd8")
cell_label <- c("% of CD8+ T cells")

for (i in 1:length(cell_class)){
  print(cell_class[i])
  pdf(file.path(save_dir, paste0(cell_class[i], "_foxp3_class.pdf")))
  
  p <- ggboxplot(subset(df_cl_l_3, !is.na(foxp3_class)), x = 'foxp3_class', y = cell_class[i], add = 'jitter', color = 'foxp3_class', palette = c("firebrick4", "tan3"), facet.by = "Hotspots") + 
    stat_compare_means(label = "p.format", label.x = 1.3, paired = F) + #, label.y = 1
    xlab('') + ylab(cell_label[i]) +
    cus_theme
  
  print(p)
  dev.off()
  
}

p.adjust(c(0.016, 0.0048, 0.024, 8.3*10^-6, 0.42, 0.11), "BH")
p.adjust(c(0.0034, 2.2*10^-7, 0.23, 0.7), "BH")
p.adjust(c(1.4*10^-5, 0.06), "BH")

```

#Supplementary

## Table. S3 Prognostic assessment of clinical parameters across 100 random splits of discovery and validation cohorts in TCGA LUSC patients.
```{r}
#====================S_intra/immune====================#

p_value_all <- c()
th_value <- c()
iteration_no <- 100
set.seed(05112022)

LUSCmean_ <- LUSCmean

for (i in 1:iteration_no){
LUSCdisc <- sample_n(LUSCmean_, floor(nrow(LUSCmean_)/2))
LUSCvalid <- subset(LUSCmean_, !(bcr_patient_barcode %in% LUSCdisc$bcr_patient_barcode))

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("fi"))

th_value <- c(th_value, res.cutLUSCd$cutpoint$cutpoint)

res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "fiC", "fi", res.cutLUSCd)

sur_sta <- survdiff(Surv(time, event)~fiC, data = LUSCvalid)

p_value_all <- c(p_value_all, sur_sta$pvalue)
  
}
#Thresholds
mean(th_value)
sd(th_value)

#percentage of iterations with p<0.05
mean(p_value_all)
sd(p_value_all)
length(p_value_all[p_value_all <0.05])/iteration_no

#confidence interval
l.model <- lm(p_value_all ~ 1)
confint(l.model, level=0.95)

#====================stages====================#
p_value_all <- c()
th_value <- c()
iteration_no <- 100
set.seed(05112022)

LUSCmean_ <- LUSCmean
LUSCmean_ <- subset(LUSCmean_, !is.na(stage_merge))

for (i in 1:iteration_no){
LUSCdisc <- sample_n(LUSCmean_, floor(nrow(LUSCmean_)/2))
LUSCvalid <- subset(LUSCmean_, !(bcr_patient_barcode %in% LUSCdisc$bcr_patient_barcode))

sur_sta <- survdiff(Surv(time, event)~stage_merge, data = LUSCvalid)

p_value_all <- c(p_value_all, sur_sta$pvalue)
}

#percentage of iterations with p<0.05
mean(p_value_all)
sd(p_value_all)
length(p_value_all[p_value_all <0.05])/iteration_no

l.model <- lm(p_value_all ~ 1)
confint(l.model, level=0.95)

#====================Age====================#
p_value_all <- c()
th_value <- c()
iteration_no <- 100
set.seed(05112022)

LUSCmean_ <- LUSCmean
LUSCmean_ <- subset(LUSCmean_, !is.na(age))

for (i in 1:iteration_no){
LUSCdisc <- sample_n(LUSCmean_, floor(nrow(LUSCmean_)/2))
LUSCvalid <- subset(LUSCmean_, !(bcr_patient_barcode %in% LUSCdisc$bcr_patient_barcode))

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("age"))

th_value <- c(th_value, res.cutLUSCd$cutpoint$cutpoint)

res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "ageC", "age", res.cutLUSCd)

sur_sta <- survdiff(Surv(time, event)~ageC, data = LUSCvalid)

p_value_all <- c(p_value_all, sur_sta$pvalue)
  
}
#Thresholds
mean(th_value)
sd(th_value)

#percentage of iterations with p<0.05
mean(p_value_all)
sd(p_value_all)
length(p_value_all[p_value_all <0.05])/iteration_no

#confidence interval
l.model <- lm(p_value_all ~ 1)
confint(l.model, level=0.95)

#====================Pack years====================#
p_value_all <- c()
th_value <- c()
iteration_no <- 100
set.seed(05112022)

LUSCmean_ <- LUSCmean
LUSCmean_ <- subset(LUSCmean_, !is.na(pack_years))

for (i in 1:iteration_no){
LUSCdisc <- sample_n(LUSCmean_, floor(nrow(LUSCmean_)/2))
LUSCvalid <- subset(LUSCmean_, !(bcr_patient_barcode %in% LUSCdisc$bcr_patient_barcode))

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("pack_years"))

th_value <- c(th_value, res.cutLUSCd$cutpoint$cutpoint)

res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "pack_yearsC", "pack_years", res.cutLUSCd)

sur_sta <- survdiff(Surv(time, event)~pack_yearsC, data = LUSCvalid)

p_value_all <- c(p_value_all, sur_sta$pvalue)
  
}
#Thresholds
mean(th_value)
sd(th_value)

#percentage of iterations with p<0.05
mean(p_value_all)
sd(p_value_all)
length(p_value_all[p_value_all <0.05])/iteration_no

#confidence interval
l.model <- lm(p_value_all ~ 1)
confint(l.model, level=0.95)

#====================TMB====================#
p_value_all <- c()
th_value <- c()
iteration_no <- 100
set.seed(05112022)

LUSCmean_ <- LUSCmean
LUSCmean_ <- subset(LUSCmean_, !is.na(total_perMB))

for (i in 1:iteration_no){
LUSCdisc <- sample_n(LUSCmean_, floor(nrow(LUSCmean_)/2))
LUSCvalid <- subset(LUSCmean_, !(bcr_patient_barcode %in% LUSCdisc$bcr_patient_barcode))

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("total_perMB"))

th_value <- c(th_value, res.cutLUSCd$cutpoint$cutpoint)

res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCvalid <- stratify_df(LUSCvalid, "total_perMBC", "total_perMB", res.cutLUSCd)

sur_sta <- survdiff(Surv(time, event)~total_perMBC, data = LUSCvalid)

p_value_all <- c(p_value_all, sur_sta$pvalue)
  
}
#Thresholds
mean(th_value)
sd(th_value)

#percentage of iterations with p<0.05
mean(p_value_all)
sd(p_value_all)
length(p_value_all[p_value_all <0.05])/iteration_no

#confidence interval
l.model <- lm(p_value_all ~ 1)
confint(l.model, level=0.95)
```

## Fig. S1 Validation of H&E image analysis for NSCLC + Fig. S9D
```{r}

#merges
LUADtumor <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUADmean, LUAD.TCGAPathScores, estLUAD2, absolute2))
LUSCtumor <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUSCmean, LUSC.TCGAPathScores, estLUSC2, absolute2))
LUADimmune <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUADmean, estLUAD2, BoLi2, Davoli))
LUSCimmune <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUSCmean, estLUSC2, BoLi2, Davoli))
LUADstroma <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUADmean, LUAD.TCGAPathScores, estLUAD2))
LUSCstroma <- Reduce(function(x, y) merge(x, y, by ="bcr_patient_barcode", all.x=TRUE), list(LUSCmean, LUSC.TCGAPathScores, estLUSC2))
#prepare
rownames(LUADtumor) <- LUADtumor$bcr_patient_barcode
rownames(LUSCtumor) <- LUSCtumor$bcr_patient_barcode
rownames(LUADimmune) <- LUADimmune$bcr_patient_barcode
rownames(LUSCimmune) <- LUSCimmune$bcr_patient_barcode
rownames(LUADstroma) <- LUADstroma$bcr_patient_barcode
rownames(LUSCstroma) <- LUSCstroma$bcr_patient_barcode

LUADtumor <- subset(LUADtumor, select = c(tumour_per, TCGA_pathTumorPer, ARAN_ESTIMATE, ARAN_ABSOLUTE) )
LUSCtumor <- subset(LUSCtumor, select = c(tumour_per, TCGA_pathTumorPer, ARAN_ESTIMATE, ARAN_ABSOLUTE) )

LUSCimmunefiMCPONLY <-subset(LUSCimmune, select = c(fi, MCPcounter_Tcells, MCPcounter_CD8Tcells, MCPcounter_Cytotoxiclymphocytes, MCPcounter_NKcells, MCPcounter_Blineage, 
                                                    MCPcounter_Monocyticlineage, MCPcounter_Myeloiddendriticcells, MCPcounter_Neutrophils, MCPcounter_Endothelialcells))

LUSCimmunefi <-subset(LUSCimmune, select = c(fi, DAVOLI_TotNofMutations.in.exons, DAVOLI_TP53Mutation, DAVOLI_CellCycle.Signature.Score, DAVOLI_SCNA.Level, DAVOLI_Chrom.SCNA.Level, 
                                             DAVOLI_Arm.SCNA.Level, DAVOLI_Focal.SCNA.Level, DAVOLI_Chrom.Arm.SCNA.Level, DAVOLI_SCNA.Level.normalized.by.size, 
                                             MCPcounter_Tcells, MCPcounter_CD8Tcells, MCPcounter_Cytotoxiclymphocytes, MCPcounter_NKcells, MCPcounter_Blineage, 
                                             MCPcounter_Monocyticlineage, MCPcounter_Myeloiddendriticcells, MCPcounter_Neutrophils, MCPcounter_Endothelialcells))

LUADimmune <-subset(LUADimmune, select = c(fi, lym_per, ESTIMATE_ImmuneScore, MCPcounter_immune_sum, TIMER_B_cell, TIMER_T_cell.CD4, TIMER_T_cell.CD8, 
                                           TIMER_Neutrophil, TIMER_Macrophage, TIMER_DC, DAVOLI_TotNofMutations.in.exons, DAVOLI_TP53Mutation, 
                                           DAVOLI_Immune.Signature.Score, DAVOLI_CellCycle.Signature.Score, DAVOLI_SCNA.Level, DAVOLI_Chrom.SCNA.Level, 
                                           DAVOLI_Arm.SCNA.Level, DAVOLI_Focal.SCNA.Level, DAVOLI_Chrom.Arm.SCNA.Level, DAVOLI_SCNA.Level.normalized.by.size))
LUSCimmune <-subset(LUSCimmune, select = c(fi, lym_per, ESTIMATE_ImmuneScore, MCPcounter_immune_sum, TIMER_B_cell, TIMER_T_cell.CD4, TIMER_T_cell.CD8, 
                                           TIMER_Neutrophil, TIMER_Macrophage, TIMER_DC, DAVOLI_TotNofMutations.in.exons, DAVOLI_TP53Mutation, 
                                           DAVOLI_Immune.Signature.Score, DAVOLI_CellCycle.Signature.Score, DAVOLI_SCNA.Level, DAVOLI_Chrom.SCNA.Level, 
                                           DAVOLI_Arm.SCNA.Level, DAVOLI_Focal.SCNA.Level, DAVOLI_Chrom.Arm.SCNA.Level, DAVOLI_SCNA.Level.normalized.by.size))


LUADstroma <-subset(LUADstroma, select = c(stromal_per, TCGA_pathStromalPer, ESTIMATE_StromalScore, MCPcounter_stromal_sum))
LUSCstroma <-subset(LUSCstroma, select = c(stromal_per, TCGA_pathStromalPer, ESTIMATE_StromalScore, MCPcounter_stromal_sum))

tumour <- rbind(LUADtumor, LUSCtumor)
stroma <- rbind(LUADstroma, LUSCstroma)
immune <- rbind(LUADimmune, LUSCimmune)
immune <- subset(immune, select = -c(fi, DAVOLI_TotNofMutations.in.exons, DAVOLI_TP53Mutation, DAVOLI_SCNA.Level, DAVOLI_SCNA.Level, DAVOLI_Chrom.SCNA.Level, 
                                     DAVOLI_Arm.SCNA.Level, DAVOLI_Focal.SCNA.Level, DAVOLI_Chrom.Arm.SCNA.Level, DAVOLI_SCNA.Level.normalized.by.size, DAVOLI_CellCycle.Signature.Score))

tumour <- na.omit(tumour)
stroma <- na.omit(stroma)
immune <- na.omit(immune)
LUADtumor <- na.omit(LUADtumor)
LUSCtumor <- na.omit(LUSCtumor)
LUADimmune <- na.omit(LUADimmune)
LUSCimmune <- na.omit(LUSCimmune)
LUADstroma <- na.omit(LUADstroma)
LUSCstroma <- na.omit(LUSCstroma)
LUSCimmunefi <- na.omit(LUSCimmunefi)
LUSCimmunefiMCPONLY <- na.omit(LUSCimmunefiMCPONLY)

### Corr maps
corTumor <- round(cor(tumour),2)
corStroma <- round(cor(stroma),2)
corImmune <- round(cor(immune),2)
corLUADtumor <- round(cor(LUADtumor),2)
corLUSCtumor <- round(cor(LUSCtumor),2)
corLUADimmune <- round(cor(LUADimmune),2)
corLUSCimmune <- round(cor(LUSCimmune),2)
corLUADstroma <- round(cor(LUADstroma),2)
corLUSCstroma <- round(cor(LUSCstroma),2)
corLUSCimmunefi <- round(cor(LUSCimmunefi),2)
corLUSCimmunefiMCPONLY <- round(cor(LUSCimmunefiMCPONLY),2)

cor.test(LUSCimmunefiMCPONLY$fi, LUSCimmunefiMCPONLY$MCPcounter_CD8Tcells)

ggcorrplot(corTumor, method = "circle", title = "TCGA NSCLC - cancer cell, n=401")
ggcorrplot(corStroma, method = "circle", title = "TCGA NSCLC - stroma cell, n=880")
ggcorrplot(corImmune, method = "circle", title = "TCGA NSCLC - immune cell, n=689")

#Fig. S9D
ggcorrplot(corLUSCimmunefiMCPONLY, method = "circle", title = "TCGA LUSC - immune hotspot, n=463")

```


## Fig. S2 DEGs related to Sintra/immune in TCGA LUAD
```{r}
#Compare gene expr of fi high low in LUAD
variable_pair <- c("High", "Low")
variable_col <- "fi_class"
#Volcano plot
TCGAVisualize_volcano(x = dataDEGs_LUAD$logFC,
                    y = dataDEGs_LUAD$FDR,
                    filename = file.path(save_dir, paste0(variable_col, "_", variable_pair[1], "_", variable_pair[2], "_volcano_LUAD.pdf")),
                    x.cut = 3,
                    y.cut = 10^-5,
                    names = rownames(dataDEGs_LUAD),
                    color = c("black","red","darkgreen"),
                    names.size = 2,
                    xlab = " Gene expression fold change (Log2)",
                    legend = "State",
                    highlight = c("CD19", "CD79B", "MS4A1", "CD20", "CXCR5", "FCER2"),
                    title = paste0("Volcano plot (", variable_col, ", ", variable_pair[1], " vs ", variable_pair[2], ")"),
                    show.names = "both",
                    width = 10)
#EA plot
TCGAvisualize_EAbarplot(tf = rownames(ansEA_LUAD$ResBP), 
                    GOBPTab = ansEA_LUAD$ResBP,
                    GOCCTab = ansEA_LUAD$ResCC,
                    GOMFTab = ansEA_LUAD$ResMF,
                    PathTab = ansEA_LUAD$ResPat,
                    nRGTab = Genelist_LUAD, 
                    nBar = 10,
                    filename=file.path(save_dir, paste0(variable_col, "_", variable_pair[1], "_", variable_pair[2], "_logFc_all_LUAD.pdf")))

```

## Fig. S3 Assessment of registration quality and the lymphocyte densities of serial H&E and IHC sections of 10 LUSC patients.
```{r}
#============ A ===============#
rtre$mean_rTRE <- rtre$mean_rTRE*100
ggdensity(rtre, x = 'mean_rTRE', rug = TRUE, add = 'mean', xlab = 'Mean Target Registration Error (%)', ylab = 'Density', color = "tomato2") + cus_theme

#============ B ===============#

lym_list <- c("cd4","cd8","foxp3","cd20","cd20cxcr5","cd79bCoexp")

sum_df <- data.frame(slide  = sum_valid$slide, hs = sum_valid$Hotspots, p40=sum_valid$p40, 
                     cell.count.c = sum_valid$cell.count.c, cell.count.l =sum_valid$cell.count.l,
                     lym_ihc = rowSums(sum_valid[,lym_list]), x.s= sum_valid$x.s, y.s = sum_valid$y.s)

df <- subset(sum_df, hs %in% c('Cancer-immune','Immune'))  %>% 
  group_by(slide,hs) %>% 
  summarise(p40 = sum(p40,na.rm=TRUE), cell.count.c = sum(cell.count.c,na.rm=TRUE), 
            cell.count.l = sum(cell.count.l, na.rm=TRUE),
            lym_ihc = sum(lym_ihc, na.rm=TRUE), area = paste(x.s, y.s) %>% unique %>% length())

df[,setdiff(colnames(df), c("slide","hs"))] <- df[,setdiff(colnames(df), c("slide","hs"))] / df$area / (50*50) * 10^6
#levels(df$hs) <- c('Cancer','Cancer-immune','Else','Immune')
colnames(df)[which(colnames(df)=='hs')] <- 'Hotspots'

#cancer
scatter_cancer <- ggscatter(df, x = 'cell.count.c', y = 'p40', add = "reg.line",                
          conf.int = TRUE, color='Hotspots',  palette = c("#EE0000","#00BFFF"),              
          add.params = list(color = "gray40",
                            fill = "lightgray"), 
          ylab = expression("P40"^"+"~"cell / "~mm^2),
          xlab = expression("H&E-based cancer cell / "~mm^2)) + stat_cor(method = "pearson") + cus_theme + coord_fixed(ratio = 1)

#lymphocyte
scatter_lym <- ggscatter(df, x = 'cell.count.l', y = 'lym_ihc', add = "reg.line",                     
          conf.int = TRUE,  color = 'Hotspots', palette = c("#EE0000","#00BFFF"),              
          add.params = list(color = "gray40",
                            fill = "lightgray"), 
          xlab = expression("H&E-based lymphocyte cell / "~mm^2),
          ylab = expression("IHC-based lymphocyte cell / "~mm^2)) + stat_cor(method = "pearson") + cus_theme + coord_fixed(ratio = 0.6)

ggarrange(scatter_cancer, scatter_lym, ncol = 2)

#============ C ===============#
df <- sum_valid %>% group_by(slide, Hotspots) %>% summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
                                                                  foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE),
                                                                  cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE))


df_sum <- sum_valid %>% group_by(slide) %>% summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
                                                                      foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE),
                                                                      cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE))


df <- merge(df, df_sum, by = 'slide')


df[,3:8] <- df[,3:8]/df[,9:14] * 100
df_m <- melt(df[,1:8], id.vars = c("Hotspots", "slide"))
df_m$Hotspots <- factor(df_m$Hotspots, levels = c("Cancer-immune", "Immune", "Cancer", "Else"))


my_comparison<- list(c("Cancer-immune", "Immune"), c("Cancer-immune", "Cancer"))
p <- ggboxplot(df_m, x = "variable", y = "value", color = "Hotspots", add = "jitter",
          palette = c("#A0C588","#636AA8", '#C1534E', '#E4C1C1')) + stat_compare_means(comparisons = my_comparison)

df_compare <- compare_means(value ~ Hotspots, p.adjust.method = "BH", method='wilcox.test', paired = T,
              group.by = "variable",
              data =  df_m) %>%
  mutate( p.adj = format.pval(p.adj, digits = 2),
         p_adj_star = ifelse(as.numeric(gsub("< ", "", p.adj)) <0.0001, '****', 
                             ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.001, '***', 
                                    ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.01, '**', 
                                           ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.05, '*', '')))))  %>%
  mutate(line_color = ifelse(p_adj_star == "", "white", "black"))

#============ D ===============#

df <- sum_valid %>% group_by(slide, Hotspots) %>% summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE), cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE))

df_sum <- sum_valid %>% group_by(slide) %>% summarise(cd8 = sum(cd8, na.rm = TRUE), cd4 = sum(cd4, na.rm = TRUE),
foxp3 = sum(foxp3, na.rm=TRUE), cd20 = sum(cd20, na.rm=TRUE), cd20cxcr5 = sum(cd20cxcr5, na.rm = TRUE), cd79bCoexp = sum(cd79bCoexp, na.rm=TRUE))


df <- merge(df, df_sum, by = 'slide')

df[,3:8] <- df[,3:8]/rowSums(df[,9:14]) * 100
df_m <- melt(df[,1:8], id.vars = c("Hotspots", "slide"))
df_m$Hotspots <- factor(df_m$Hotspots, levels = c("Cancer-immune", "Immune", "Cancer", "Else"))


my_comparison<- list(c("Cancer-immune", "Immune"), c("Cancer-immune", "Cancer"))
p <- ggboxplot(df_m, x = "variable", y = "value", color = "Hotspots", add = "jitter",
          palette = c("#A0C588","#636AA8", '#C1534E', '#E4C1C1')) + stat_compare_means(comparisons = my_comparison)

df_compare <- compare_means(value ~ Hotspots, p.adjust.method = "BH", method='wilcox.test', paired = T,
              group.by = "variable",
              data =  df_m) %>%
  mutate( p.adj = format.pval(p.adj, digits = 2),
         p_adj_star = ifelse(as.numeric(gsub("< ", "", p.adj)) <0.0001, '****', 
                             ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.001, '***', 
                                    ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.01, '**', 
                                           ifelse(as.numeric(gsub("< ", "", p.adj)) < 0.05, '*', '')))))  %>%
  mutate(line_color = ifelse(p_adj_star == "", "white", "black"))

```

## Fig. S4 Prognostic significance of Sintra/cancer and Sintra/tissue in TCGA LUSC patients.
```{r}
LUSCdisc <- subset(LUSCmean, cohort == "disc")
#============ A, C ===============#
#Sintra/cancer
res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("fc"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCmean_ <- stratify_df(LUSCmean, "fcC", "fc", res.cutLUSCd)

##survival plot
fit <- survfit(Surv(time, event)~fcC, data = LUSCmean_)

ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

fitd <- coxph(Surv(time, event)~fcC+age+stage_merge+pack_years, data = LUSCmean_)
ggforest(fitd, main = "TCGA LUSC: discovery + validation cohort")

#============ B, D ===============#
#Sintra/tissue
res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("ft"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCmean_ <- stratify_df(LUSCmean, "ftC", "ft", res.cutLUSCd)

##survival plot
fit <- survfit(Surv(time, event)~ftC, data = LUSCmean_)

ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

fitd <- coxph(Surv(time, event)~ftC+age+stage_merge+pack_years, data = LUSCmean_)
ggforest(fitd, main = "TCGA LUSC: discovery + validation cohort")

```

## Fig. S5 Prognostic significance of Sintra/peri in TCGA NSCLC patients with different smoking histories.
```{r}
#LUAD
p_sur_luad_1 <- plt_sur_by_smokingClass(LUADmean, "1", "Current reformed smoker for <= 15 years")
p_sur_luad_2 <- plt_sur_by_smokingClass(LUADmean, "2", "Current reformed smoker for > 15 years")
p_sur_luad_3 <- plt_sur_by_smokingClass(LUADmean, "3", "Current smoker")
p_sur_luad_4 <- plt_sur_by_smokingClass(LUADmean, "4", "Lifelong non-smoker")

LUADmean_1 <- subset(LUADmean, smokingClass %in% c('1','2','3','4'))
LUADmean_1<- subset(LUADmean_1, time>=0)

LUADdisc_1 <- LUADmean_1[1:round(nrow(LUADmean_1)/2) ,]

res.cutLUADd <- surv_cutpoint(LUADdisc_1, time = "time", event = "event",
                              variables = c("fi"))
res.catLUADd <- surv_categorize(res.cutLUADd)

LUADmean_1 <- stratify_df(LUADmean_1, "fiC", "fi", res.cutLUADd)

fita <- coxph(Surv(time, event)~fiC+stage_merge+smokingClass, data = LUADmean_1)
ggforest(fita, data = NULL, main = "", fontsize = 0.7, refLabel = "", noDigits = 2)

#LUSC
p_sur_lusc_1 <- plt_sur_by_smokingClass(LUSCmean, "1", "Current reformed smoker for <= 15 years")
p_sur_lusc_2 <- plt_sur_by_smokingClass(LUSCmean, "2", "Current reformed smoker for > 15 years")
p_sur_lusc_3 <- plt_sur_by_smokingClass(LUSCmean, "3", "Current smoker")
p_sur_lusc_4 <- plt_sur_by_smokingClass(LUSCmean, "4", "Lifelong non-smoker")

LUSCmean_1 <- subset(LUSCmean, smokingClass %in% c('1','2','3','4'))
LUSCmean_1<- subset(LUSCmean_1, time>=0)

LUSCdisc_1 <- LUSCmean_1[1:round(nrow(LUSCmean_1)/2) ,]

res.cutLUSCd <- surv_cutpoint(LUSCdisc_1, time = "time", event = "event",
                              variables = c("fi"))

LUSCmean_1 <- stratify_df(LUSCmean_1, "fiC", "fi", res.cutLUSCd)

fita <- coxph(Surv(time, event)~fiC+stage_merge+smokingClass, data = LUSCmean_1)
ggforest(fita, data = NULL, main = "", fontsize = 0.7, refLabel = "", noDigits = 2)

```

## Fig. S6 Prognostic significance of three immune scores in TCGA LUAD patients
```{r}
LUADdisc <- subset(LUADmean, cohort == "disc")
LUADvalid <- subset(LUADmean, cohort == "valid")

#S_intra/immune
#########################
#Optimize threshold for the discovery set
res.cutLUADd <- surv_cutpoint(LUADdisc, time = "time", event = "event",
                              variables = c("fi"))
res.catLUADd <- surv_categorize(res.cutLUADd)

#disc
LUADdisc <- stratify_df(LUADdisc, "fiC", "fi", res.cutLUADd)

fit <- survfit(Surv(time, event)~fiC, data = LUADdisc)

ggsurvplot(fit, data = LUADdisc, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#valid
LUADvalid <- stratify_df(LUADvalid, "fiC", "fi", res.cutLUADd)

fit <- survfit(Surv(time, event)~fiC, data = LUADvalid)

ggsurvplot(fit, data = LUADvalid, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#combined

LUADmean <- stratify_df(LUADmean, "fiC", "fi", res.cutLUADd)

fit <- survfit(Surv(time, event)~fiC, data = LUADmean)

ggsurvplot(fit, data = LUADmean, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
#########################

#S_intra/cancer
#########################
#Optimize threshold for the discovery set
res.cutLUADd <- surv_cutpoint(LUADdisc, time = "time", event = "event",
                              variables = c("fc"))
res.catLUADd <- surv_categorize(res.cutLUADd)

#disc
LUADdisc <- stratify_df(LUADdisc, "fcC", "fc", res.cutLUADd)

fit <- survfit(Surv(time, event)~fcC, data = LUADdisc)

ggsurvplot(fit, data = LUADdisc, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#valid
LUADvalid <- stratify_df(LUADvalid, "fcC", "fc", res.cutLUADd)

fit <- survfit(Surv(time, event)~fcC, data = LUADvalid)

ggsurvplot(fit, data = LUADvalid, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#combined

LUADmean <- stratify_df(LUADmean, "fcC", "fc", res.cutLUADd)

fit <- survfit(Surv(time, event)~fcC, data = LUADmean)

ggsurvplot(fit, data = LUADmean, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
#########################

#S_intra/tissue
#########################
#Optimize threshold for the discovery set
res.cutLUADd <- surv_cutpoint(LUADdisc, time = "time", event = "event",
                              variables = c("ft"))
res.catLUADd <- surv_categorize(res.cutLUADd)

#disc
LUADdisc <- stratify_df(LUADdisc, "ftC", "ft", res.cutLUADd)

fit <- survfit(Surv(time, event)~ftC, data = LUADdisc)

ggsurvplot(fit, data = LUADdisc, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#valid
LUADvalid <- stratify_df(LUADvalid, "ftC", "ft", res.cutLUADd)

fit <- survfit(Surv(time, event)~ftC, data = LUADvalid)

ggsurvplot(fit, data = LUADvalid, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

#combined

LUADmean <- stratify_df(LUADmean, "ftC", "ft", res.cutLUADd)

fit <- survfit(Surv(time, event)~ftC, data = LUADmean)

ggsurvplot(fit, data = LUADmean, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
#########################
```

## Fig. S7 Survival test of lymphocyte abundance in TCGA LUSC and LUAD cohort.
```{r}
LUSCdisc <- subset(LUSCmean, cohort == "disc")

#============ A, C ===============#

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event",
                              variables = c("lym_per"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUSCmean_ <- stratify_df(LUSCmean, "lym_perC", "lym_per", res.cutLUSCd)

##survival plot
fit <- survfit(Surv(time, event)~lym_perC, data = LUSCmean_)

ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

fitd <- coxph(Surv(time, event)~lym_perC+age+stage_merge+pack_years, data = LUSCmean_)
ggforest(fitd, main = "TCGA LUSC: discovery + validation cohort")

#============ B, D ===============#
LUADdisc <- subset(LUADmean, cohort == "disc")

res.cutLUSCd <- surv_cutpoint(LUADdisc, time = "time", event = "event",
                              variables = c("lym_per"))
res.catLUSCd <- surv_categorize(res.cutLUSCd)

LUADmean_ <- stratify_df(LUADmean, "lym_perC", "lym_per", res.cutLUSCd)

##survival plot
fit <- survfit(Surv(time, event)~lym_perC, data = LUADmean_)

ggsurvplot(fit, data = LUADmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))

fitd <- coxph(Surv(time, event)~lym_perC+age+stage_merge+pack_years, data = LUADmean_)
ggforest(fitd, main = "TCGA LUAD: discovery + validation cohort")

```

## Fig. S9 Immunological phenotype of immune hotspots orchestrated by infiltrating B cells and T regulatory cells.
```{r}
#============ A ===============#

#exhaustedB signature 
grep("CD27", rownames(dataFiltT), value = TRUE)
exhaustedB <- as.data.frame( dataFiltT[c("CD19", "CD4", "CD69", "CD27", "MS4A1", "FOXP3", 
                                         "CD8A", "CD8B" ,"CD80" ,"CD86", "CD81" ,"CD82", "CD84", "CD83", "CR2" ),])
exhaustedB <- as.data.frame(t(exhaustedB))
exhaustedB$id2 <- rownames(exhaustedB)

exhaustedB <- merge(exhaustedB, col, by = "id2")
#take out the outlier in row 381 TCGA-85-A4PA-01A-11R-A24Z-07
#exhaustedB <- exhaustedB[-c(381),]

ggscatter(exhaustedB, x = "FOXP3", y = "CD27", color = "condition",
          add = "reg.line", conf.int = TRUE, size = 2,
          add.params = list(color = "blue",
                            fill = "lightgray"), title = "TCGA LUSC (n = 466)",
          cor.coef = TRUE, cor.method = "spearman")
my_comparisons <- list( c("high", "low"))

ggboxplot(exhaustedB, x = "condition", y = "FOXP3", add = "jitter", color = "condition", title = "TCGA LUSC (n=466)",
          palette = c("#EE0000", "#00BFFF")) +stat_compare_means(comparisons = my_comparisons)
ggboxplot(exhaustedB, x = "condition", y = "CD27", add = "jitter", color = "condition", title = "TCGA LUSC (n=466)",
          palette = c("#EE0000", "#00BFFF")) +stat_compare_means(comparisons = my_comparisons)


#take out the outlier in row 381 TCGA-85-A4PA-01A-11R-A24Z-07
exhaustedBOutlier <- exhaustedB[-c(381),]
exhaustedBOutlier <- exhaustedBOutlier[order(exhaustedBOutlier$condition),]  #order for the heatmap!
annot <- subset(exhaustedBOutlier, select = c(id2, condition))
rownames(annot) <- annot$id2
annot <- subset(annot, select = -c(id2))

mapBcell <- subset(exhaustedBOutlier, select = -c(bcr_patient_barcode, id, condition, 
                                                  CD8B, CD80, CD86, CD81, CD82, CD84, CD83))

corBcell <- subset(mapBcell, select = -c(id2))
rownames(mapBcell) <- mapBcell$id2
mapBcell <- subset(mapBcell, select = -c(id2))
mapBcell <- t(mapBcell)
condition = c("#EE0000", "#00BFFF")
names(condition) = c("high", "low")
ann_colors = list(condition = condition)

#============ B ===============#
mapBcelllog <- log10(mapBcell+1)
pheatmap(mapBcelllog, annotation = annot,annotation_legend = T, annotation_colors = ann_colors, show_colnames= F, 
         gaps_col =  3, cluster_rows=T, cluster_cols=T, drop_levels = F, border_color = NA)

grid.ls(grid.force()) # "col_annotation" looks like it's the one to edit
# grid.gedit("col_annotation", gp = gpar(col="grey70"))

#============ C ===============#
matcorBcell <- round(cor(corBcell),2)
ggcorrplot(matcorBcell, method = "square", hc.order = TRUE, type = "upper", lab=TRUE,
           outline.col = "white", title = "TCGA LUSC, n=462")

```

##Fig. S10 The association between S_(intra/immune) and TMB, PD1, PD-L1, and neoantigens in LUSC
- ======TMB======
- Download maf files and clinical data for TCGA-LUSC from the [GDC portal](https://portal.gdc.cancer.gov/repository)
- Select the authorized files (549 from 490 cases)
- Combine all maf files into one: bash move_maf.sh maf_dir LUSC_549.maf
- Format the 'Tumor_Sample_Barcode' column in LUSC_549.maf to Project-TSS-Participant: python tcga_format.py LUSC_549.maf > LUSC_549_sort.maf
- For the clinical.tsv, change column name 'submitted_id' to 'Tumor_Sample_Barcode'
- Analyse with [maftools](https://www.bioconductor.org/packages/devel/bioc/vignettes/maftools/inst/doc/maftools.html)
- Merge TMB column with LUSCmean dataset
```{r}
#============ A ===============#
lusc.mutload_merge <- merge(lusc.mutload, col, by.x = "Tumor_Sample_Barcode", by.y = "bcr_patient_barcode", all.y=TRUE)
lusc.mutload_merge <- subset(lusc.mutload_merge, !is.na(total_perMB_log))
lusc.mutload <- lusc.mutload[order(lusc.mutload$Tumor_Sample_Barcode),]

#plot
my_comparisons <- list( c("high", "low"))

pdf(file.path(save_dir, "TMB_high_low.pdf"), width = 6, height = 6)
p <- ggboxplot(lusc.mutload_merge, x = "condition", y = "total_perMB", add = "jitter", color = "condition", title = paste0("TCGA LUSC (n=", nrow(lusc.mutload_merge), ")"),
          palette = c("#EE0000", "#00BFFF")) +stat_compare_means(comparisons = my_comparisons) +
          cus_theme +
          xlab("S_intra/immune") + ylab('Somatic mutation frequency (/Mb)')
print(p)
dev.off()

#============ B, C, D ===============#
load("data/survival_LUSC_LUAD_v3.RData")

LUSCmean_ <- subset(LUSCmean, !is.na(total_perMB))

LUSCdisc <- subset(LUSCmean_, cohort == 'disc')
LUSCvalid <- subset(LUSCmean_, cohort == 'valid')

res.cutLUSCd <- surv_cutpoint(LUSCdisc, time = "time", event = "event", minprop = 0.2,
                             variables = c("total_perMB", "fi"))

LUSCmean_ <- stratify_df(LUSCmean_, "fiC", "fi", res.cutLUSCd)
LUSCmean_ <- stratify_df(LUSCmean_, "total_perMBC", "total_perMB", res.cutLUSCd)

fit <- survfit(Surv(time, event)~total_perMBC + fiC, data = LUSCmean_)

pdf(file.path(save_dir, "sur_TMB_fi_optimal_disc.pdf"), width = 11, height = 8, onefile = F)
ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           #palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
dev.off()

#TMB survival on its own

LUSCmean_$total_perMBC <- factor(LUSCmean_$total_perMBC, levels = c("High", "Low"))
fit <- survfit(Surv(time, event)~total_perMBC, data = LUSCmean_)

pdf(file.path(save_dir, "sur_TMB_fi.pdf"), width = 11, height = 8, onefile = F)
ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
dev.off()

#TMB High + Sintra/immune low

LUSCmean_$Stratification <- ifelse(LUSCmean_$total_perMB > res.cutLUSCd$total_perMB$estimate[[1]] & LUSCmean_$fi <= res.cutLUSCd$fi$estimate[[1]], "High TMB low S_intra/immune", "Others")

fit <- survfit(Surv(time, event)~Stratification, data = LUSCmean_)

pdf(file.path(save_dir, "sur_TMB_fi_merge.pdf"), width = 13, height = 8, onefile = F)
ggsurvplot(fit, data = LUSCmean_, risk.table = T,  break.time.by = 500, 
           ylab = "Overall Survival",
           xlab = "Time (Days)",
           palette = c("#00BFFF", "#EE0000"),
           pval = T, pval.coord = c(0.1, 0.1))
dev.off()


#Forest plot

LUSCmean_$total_perMBC <- factor(LUSCmean_$total_perMBC, levels = c("High", "Low"))

fita <- coxph(Surv(time, event)~fiC + total_perMBC + age + stage_merge + pack_years, data = LUSCmean_)

pdf(file.path(save_dir, "sur_TMB_optimal_multi_fi_disc_v1.pdf"), width = 8, height = 8, onefile = F)
ggforest(fita, data = LUSCmean_, main = "", fontsize = 0.7, refLabel = "", noDigits = 2)
dev.off()

#============ E, F ===============#

#PDL1 https://www.ncbi.nlm.nih.gov/gene/29126
grep("CD274", rownames(dataFiltT), value = TRUE)
#PD1 https://www.ncbi.nlm.nih.gov/gene/5133
grep("PDCD1", rownames(dataFiltT), value = TRUE)

pd1_pdl1 <- as.data.frame( dataFiltT[c("PDCD1", "CD274" ),])
pd1_pdl1 <- as.data.frame(t(pd1_pdl1))
pd1_pdl1$id2 <- rownames(pd1_pdl1)

pd1_pdl1 <- merge(pd1_pdl1, col, by = "id2")

my_comparisons <- list( c("high", "low"))

#PD1, p = 0.19
pdf(file.path(save_dir, "PD1_LUSC_S_high_low.pdf"), width = 6, height = 6)
ggboxplot(pd1_pdl1, x = "condition", y = "PDCD1", add = "jitter", color = "condition", title = paste0("TCGA LUSC (n=", dim(pd1_pdl1)[1], ")"),
          palette = c("#EE0000", "#00BFFF")) + ylab('PD1 (normalized read counts)') +
          stat_compare_means(comparisons = my_comparisons)
dev.off()

#PDL1, p = 0.12
pdf(file.path(save_dir, "PDL1_LUSC_S_high_low.pdf"), width = 6, height = 6)
ggboxplot(pd1_pdl1, x = "condition", y = "CD274", add = "jitter", color = "condition", title = paste0("TCGA LUSC (n=", dim(pd1_pdl1)[1], ")"),
          palette = c("#EE0000", "#00BFFF")) + ylab('PD-L1 (normalized read counts)') +
          stat_compare_means(comparisons = my_comparisons)
dev.off()

#============ G ===============#
#Clonal neoantigens
res.cutLUSC <- surv_cutpoint(LUSCmean, time = "time", event = "event",
                              variables = c("fi"), minprop = 0.2)
res.catLUSC <- surv_categorize(res.cutLUSC)
LUSCmean_ <- LUSCmean
LUSCmean_$fi_class <- ifelse(LUSCmean_$fi > res.cutLUSC$fi$estimate, "High", "Low")

pdf(file.path(save_dir, "LUSC_neoantigen_fi.pdf"), width = 6, height = 6, onefile = F)
ggboxplot(subset(LUSCmean_,!is.na(fi_class)), x = "fi_class", y = "ClonalNeoAntigens", add = "jitter", color = "fi_class", title = paste0("TCGA LUSC (n=", dim(subset(LUSCmean_,!is.na(ClonalNeoAntigens)))[1], ")"),
          palette = c("#EE0000", "#00BFFF")) + #+ ylab('PD1 (normalized read counts)') +
          stat_compare_means()
dev.off()

```

